\documentclass{article}

\input{preamble-articles}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{Internal models of type theory}
\author{Egbert Rijke}
\date{\today}

\begin{document}

\maketitle

\tableofcontents

\begin{comment}
\section*{Introduction\phantomsection\addcontentsline{toc}{section}{Introduction}}
This paper contributes to the univalent homotopy type theory~\cite{kapulkin2012univalence,awodey2012type,pelayo2012homotopy}
program which connects and extends a number of topics such as type theory, (elementary) topos
theory~\cite{MacLaneMoerdijk,johnstone:elephant}, homotopy theory, higher category theory and higher topos
theory~\cite{rezk2010toposes,lurie2009higher}.
A key missing ingredient is a first-order (`elementary') definition of a higher topos. It has been
conjectured~\cite{awodey2012type,shulman2012univalence} that homotopy type theory can be used as the internal language, and thus may
provide an elementary definition of a higher topos; see also~\cite{LumsdaineWarren}.

A main ingredient in the theory of higher toposes is the descent property, which
is described in~\cite{rezk2010toposes,lurie2009higher}. The (weak) descent property
of an ordinary topos $\mathcal{E}$ is the property that
\begin{enumerate}
\item For any family $\{X_i\}_{i\in I}$ of objects of $\mathcal{E}$ with coproduct
$X:=\coprod_{i\in I}X_i$ and any function $f:Y\to X$ we get an isomorphism
\begin{equation*}
Y\cong\coprod_{i\in I} Y\times_X X_i
\end{equation*}
given by the canonical map.
\item Conversely, for any family $\{f_i:X_i\to Y_i\}_{i\in I}$ of maps inducing
$f:\coprod_{i\in I}X_i\to\coprod_{i\in I}Y_i$, the canonical map determines an
isomporphism
\begin{equation*}
Y_i\cong Y\times_X X_i.
\end{equation*}
\item
\item
\end{enumerate}
In model toposes, the descent property can be strengthened considerably, but
ordinary limits and colimits are to be replaced by \emph{homotopy} limits and
colimits. Since the only notion of limits and colimits in type theory is that
of homotopy limit and colimit, this is no restriction for us. The descent
property of model toposes as formulated by Rezk~\cite{rezk2010toposes} is: Let
$X:\mathbf{I}\to\mathcal{E}$ be a functor from a small category into a
model topos $\mathcal{E}$, with homotopy colimit $\bar{X}$. Then we have
\begin{enumerate}
\item Let $\bar{Y}$ be an object and let $f:\bar{Y}\to\bar{X}$ 
be any map in $\mathcal{E}$. Define
the functor $Y:\mathbf{I}\to\mathcal{E}$ by $Y(i):=Y\times_{\bar{X}} X_i$. Then
the canonical map $\typefont{hocolim}(Y)\to\bar{Y}$ is a weak equivalence.
\item Let $Y:\mathbf{I}\to\mathcal{E}$ be a functor and let $f:Y\to X$ be an
equifibered transformation from $Y$ to $X$. Define $\bar{f}:\bar{Y}\to\bar{X}$
be the homotopy colimit of $f$. Then the canonical map 
$Y_i\to \bar{Y}\times^h_{\bar{X}} X_i$ is a weak equivalence.
\end{enumerate}
These two properties are two sides of the same coin and in our type theoretical
setting it is natural to stress this. Given a diagram $D$, we will be able
to summarize the entire descent property in an equivalence between the type
$\typefont{equiFib}(D)$ of equifibered diagrams over $D$ and the type of functions
into $\typefont{colim}(D)$. We prove the descent property in 
theorem~\ref{thm:descent}.

We use universes ala Russell, universe polymorphism and typical ambiguity and hence create the illusion that $\type:\type$;
see~\cite{TheBook}. There are only a few places where we need to be explicit about universe levels.
As is common in HoTT, we will freely use the axiom of function extensionality.
Using this axiom, all limits can be constructed in type theory~\cite{Avigad:limits}.
We use higher inductive types (HITs) to construct co-limits; see~\cite{LumsdaineShulman,Shulman:HIT}.
Using the univalence axiom one can show that $\type$ is an object classifier~\cite{RijkeSpitters:Sets}.
An important feature property of $\infty$-toposes is the descent property. We prove this property in theorem~\ref{thm:descent}.

In the previous paragraph we have used informal categorical terminology such as `limits' to refer to what would be interpreted as homotopy
limits in the $\infty$-category of $\infty$-groupoids. We will continue to use this naive $\infty$-category style without any rigorous
claims to their interpretation in the model.

The paper is organized as follows. We start by defining limits and colimits of over graph. This is the most general case, as we
do not include the possibility of including higher coherence conditions.
We introduce diagrams and (co)limits in section~\ref{sec:diagram}. In section~\ref{sec:suspensions} we study suspensions.
Functions with contractible mapping cones are the same as epimorphisms, as we will see in section~\ref{sec:mapping}.
Finally, we prove the descent property.

\subsection{Notation}
A typical way we write a type in this article is:
\begin{equation}\label{eq:notation_example}
\prd{x,y:A}{p:B(x,y)}C(x,y,p)\to D(x,y,p).
\end{equation}
Assuming that $A$, $B$, $C$ and $D$ have been introduced, this type denotes a
dependent product, where the product ranges over $x,y:A$ and $p:B(x,y)$ and
the factors are the (non-dependent) function types $C(x,y,p)\to D(x,y,p)$. A
more common notation for the same type in mathematics would be
\begin{equation*}
\prod_{x,y:A}\prod_{p:B(x,y)}C(x,y,p)\to D(x,y,p),
\end{equation*}
like a big product of function spaces. Nevertheless we have chosen to use this
less common notation, because the types $A$ or $B$ can become quite complex.
% and formulas would become messy as as complex information starts `dancing
% around'. 

Another thing we notice in formula \eqref{eq:notation_example} is the use of
curly braces around $x,y:A$. This indicates that whenever we have a term
\begin{equation*}
f:\prd{x,y:A}{p:B(x,y)}C(x,y,p)\to D(x,y,p),
\end{equation*}
we will omit reference to $x$ and $y$ whenever $f$ is applied to $x,y:A$ and
$p:B(x,y)$. In other words, we will write down $f(p):C(x,y,p)\to D(x,y,p)$
instead of $f(x,y,p)$. We will do this whenever it is possible to retrieve
the terms $x$ and $y$ from the term $p$, and in many similar situations like
this. Back to formula \autoref{eq:notation_example}, it is quite possible 
that we would have used the curly braces in the introduction
of $C$ and $D$ as well, and if we had done that we would have written $C(p)$
and $D(p)$.
\end{comment}

\input{article-models-defn}

\bibliographystyle{hplain}
%\phantomsection\addcontentsline{toc}{section}{References}
\bibliography{refs}

\end{document}
